---
title: "Example Project: SEKI"
author: "Ryan Peek"
date: "*Updated: `r format(Sys.Date())`*"
output: 
  html_document:
    code_folding: hide
    highlight: pygments
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(normalizePath("../"))
library(here)
```

 > *"So what I want to do is:  
   1. correctly project my coordinates onto a map of my study site and produce a map that displays them  
   2. measure the distance of each playback point to the location of the audio recorder (which is also point 1 in both of the locations)  
   3. graph the relationship between amplitude of the recorded Mountain Chickadee (MOCH) song and distance from recorder"*

## Load Packages

Here we `source` a separate script which only loads our libraries. This is a way to have separate scripts written for specific tasks, particularly if you do these tasks over and over.

```{r libraries}
source(here("scripts/load_spatial_libraries.R"))
suppressPackageStartupMessages(library(tidyverse))
library(viridis)
```

### Load Data

These data files look weird. Here's the `head` of one:

```{bash, echo=T, eval=T}

head -n1 ../data/CENTER1_20140816_203919.Table.1.selections.txt
```

It seems they don't have any line endings...that is they basically just continue to wrap across one long line. Let's try to read them in with `read_delim` or `read_tsv`.

```{r readInDat, warning=F, message=F}

dat1 <- read_delim(file = here("data/CENTER1_20140816_203919.Table.1.selections.txt"), delim = "\t") # that worked, so we can use read_tsv

dat2a <- read_tsv(file= here("data/CENTER2_20140817_080105.Table.1.selections.txt"))
dat2b <- read_tsv(file= here("data/CENTER2_20140817_090105.Table.1.selections.txt"))

```

### Clean Data

Ok, so we have some data...but it needs some **tidying**. Good news! We know how to do this now.

 - Let's first join the `dat2a` with `dat2b` as those are from the same day...notice they have different numbers of columns though. :(
 
```{r join2a2b, warning=F, message=F}

dat2 <- full_join(dat2a, dat2b)

```
 
Ok, looks good. There's some missing data between the two, but we can follow up on that later. Let's next look at adding a column for lake 1 and lake 2, and then joining these datasets into one, even though there a few columns that are different. It will make working with the data easier since they will all be in one single file. 

```{r join1w2, warning=F, message=F}

# add lake names to each file:
dat1 <- dat1 %>% 
  rename(Dist_to_Mic = Dist_to_SM)
dat2 <- dat2 %>% 
  select(-`Freq 95% (Hz)`, -`Freq 5% (Hz)`)

names(dat1)
names(dat2)

fulldat <- full_join(dat1, dat2) # try to put the df w/ more cols as the "left" side

# make a searchable table with the DT package
DT::datatable(fulldat)


```

Great, all data in one file!  There's some missingness, but that's ok for now. 

## Make it Spatial

Ok, now we can work with our UTM data and make this a spatial data set so we can make a map. This is using the `sf` package.

```{r makeSpatial}
fulldat_sf <- st_as_sf(fulldat,
                       coords = c("UTM_X", "UTM_Y"), # can use numbers here too
                       remove = F, # don't remove these lat/lon cols from df
                       crs = 26711) # add projection, def Z11, NAD83
# These CRS didn't work: 4269, 3310, 3372, 3488, 3311, 3740, 3741, 6414, 42303, 5071, 26944
# Didn't work: 3495, 2769, 2228, 3309, 6421, 26744, 6422, 42308, 3741, 26911, 102644
# These work (sort of): 26711

# transform data here:
#fulldat_sf <- fulldat_sf %>% st_transform(crs = 4326)

```


### Make some Maps

Now we can start viewing this data...did it project correctly? Do the points look correct? Let's take a look using a few simple interactive map packages. Namely, `mapview`. We can layer multiple objects by adding them together. We'll add a few counties here for reference.

```{r mapview}

library(USAboundaries)
counties_spec <- us_counties(resolution = "low", states="CA") %>%
  filter(name %in% c("Tulare", "Inyo", "Fresno"))

mapview(counties_spec) + mapview(fulldat_sf, color="orange") 

```

Looks like one point is wonky...row 1, UTM_X is `278545` and should probably be `378545`. Likely an error from data entry? Let's change it.

```{r fixDatapoint}

fulldat <- fulldat %>% 
  mutate(UTM_X = replace(UTM_X, UTM_X==278545, 378545))

# double check:
fulldat[fulldat$UTM_X==278545,]

# re=spatialize
fulldat_sf <- st_as_sf(fulldat,
                       coords = c("UTM_X", "UTM_Y"), # can use numbers here too
                       remove = F, # don't remove these lat/lon cols from df
                       crs = 26711 ) # add NAD83/UTM Zone 11 26711

mapview(counties_spec) + mapview(fulldat_sf)
```

### Process Spatial Data

Measure from the recorder to all other points, using the `st_distance` function. This makes a matrix of every point vs. every other point. Typically lat/longs are not an ideal projection for this, so we'll likely get a warning, but the relative measurements should be fine. The first value in each of the lakes is the recorder (I believe). So we will use the first column.

```{r}
l1 <- filter(fulldat_sf, LAKE==1) # all other pts

l1_dists <- st_distance(l1)
l1_dists

l2 <- filter(fulldat_sf, LAKE==2) # all other pts

l2_dists <- st_distance(l2)
l2_dists

# to get the distance from point 1 to all other points, grab first col
dist_to_recorder <- bind_rows(tibble(rec_dists = l1_dists[,1]), tibble(rec_dists=l2_dists[,1]))
```

## PLOTS

Plot amplitude vs. distance to recorder, using the `geom_sf` function of `ggplot`. This will require installing the development version of `ggplot2` (I believe). 

```{r, eval=F, echo=T}
#install.packages("devtools")
devtools::install_github("tidyverse/ggplot2")
```


### Plot Data Spatially

```{r blankmapplot}
# bind col with original dataset:
fulldat_sf <- bind_cols(fulldat_sf, dist_to_recorder)

library(viridis)
ggplot() + 
  geom_sf(data=fulldat_sf, aes(color=rec_dists, size=`High Freq (Hz)`)) + 
  scale_color_viridis() + coord_sf() + theme_classic()

ggplot() + 
  geom_sf(data=fulldat_sf, aes(color=rec_dists, size=`Max Power (dB)`)) + 
  scale_color_viridis() + coord_sf() + theme_classic()

```

### Add GGMAP Background

Now we add some background imagery using the `ggmap` package.

```{r ggmap}

# add a ggmap background
library(ggmap) # need this package

# make sure projection is mercator so will match google
fulldat_lat_sf <- fulldat_sf %>% st_transform(crs = 4326)

# set the background map: black and white terrain
map1 <- get_map(location=c(-118.357, 36.732), crop = F,
                #color="bw",
                maptype="terrain",
                source="google",
                zoom=15)
# plot
ggmap(map1) +
  geom_sf(data=fulldat_lat_sf, aes(color=rec_dists, size=`Max Power (dB)`), inherit.aes = FALSE) + 
  scale_color_viridis() + coord_sf() + theme_classic()

# or an imagery option:
map2 <- get_map(location=c(-118.357, 36.732), crop = F,
                #color="bw",
                maptype="satellite",
                source="google",
                zoom=15)
# plot
ggmap(map2) +
  geom_sf(data=fulldat_lat_sf, aes(color=rec_dists, size=`Max Power (dB)`), inherit.aes = FALSE) + 
  scale_color_viridis() + coord_sf() + theme_classic()

```

Ok, now let's try plotting the amplitudes vs. the distance to the recorder.

```{r regressionPlot}

# high freq and by lake
ggplot() + 
    geom_smooth(data=fulldat_sf, aes(x=rec_dists, y=`High Freq (Hz)`, group=LAKE), alpha=0.3) + 
  geom_point(data=fulldat_sf, aes(x=rec_dists, y=`High Freq (Hz)`, color=rec_dists, group=LAKE)) + 
  scale_color_viridis() + theme_classic() + facet_grid(LAKE~.)

# max power by lake
ggplot() + 
  geom_smooth(data=fulldat_sf, aes(x=rec_dists, y=`Max Power (dB)`, group=LAKE), alpha=0.3) + 
  geom_point(data=fulldat_sf, aes(x=rec_dists, y=`Max Power (dB)`, color=`High Freq (Hz)`, size=rec_dists, group=LAKE)) + 
  scale_color_viridis() + theme_classic() + facet_grid(LAKE~.)

# max peak power
ggplot() + 
    geom_smooth(data=fulldat_sf, aes(x=rec_dists, y=`Peak Power (dB)`, group=LAKE), alpha=0.3) + 
  geom_point(data=fulldat_sf, aes(x=rec_dists, y=`Peak Power (dB)`, color=`High Freq (Hz)`, size=rec_dists, group=LAKE)) + 
  scale_color_viridis() + theme_classic() + facet_grid(LAKE~.)

```

