---
title: "Example Project: Suisun Marsh"
author: "Ryan Peek"
date: "*Updated: `r format(Sys.Date())`*"
output: 
  html_document:
    code_folding: hide
    highlight: pygments
    theme: sandstone
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 > *"Just collated Kyle Phillip's and my field data from Oct-recent from a 500 acre duck pond in the Suisun Marsh! Sampling was a tad discrete and inconsistent in the earlier sampling events as we were figuring things out. But we're ramping up now and understanding spatial heterogeneity is an important component to our work."*

## Load Packages

Here we `source` a separate script which only loads our libraries. This is a way to have separate scripts written for specific tasks, particularly if you do these tasks over and over.

```{r libraries, warning=FALSE}
library(here)
library(viridis)
source(here("scripts/load_spatial_libraries.R"))
suppressPackageStartupMessages(library(tidyverse))
```

### Load Data

Let's load the data and format things so we can use them.

```{r readInDat, warning=F, message=F}

dat <- read_csv(file = here("data/LucoFieldData.csv"))
glimpse(dat)

```

### Clean Data

Ok, so we have some data...but it needs some **tidying**. Good news! We know how to do this now.

 - Let's first convert our `Date` and `Time` column into a `datetime` format (`POSIXct`). 
 - Then we'll summarize and check for missingess in our data using the [`naniar`](http://naniar.njtierney.com/) package.
 
```{r format, warning=F, message=F}
library(lubridate)
dat2 <- dat %>% 
  mutate(datetime = mdy_hms(paste0(Date, " ", Time))) %>% 
  select(-Time, -Date)
glimpse(dat2)  
```
 
**Check for missingness:**

 - we can do with summary or is.na, or we can look at this visually
 
```{r naniar, eval=T, echo=T}

summary(dat2)

library(naniar)
# number missing
gg_miss_var(dat2)

# show % missing
#gg_miss_var(dat2, show_pct = T)
```


### A few simple plots

```{r}

ggplot() + 
  geom_point(data=dat2, aes(x=datetime, y=Temp_C, fill=Chlor), pch=21, size=4) +
  scale_fill_viridis() + 
  facet_grid(Pond_Slough~.)

ggplot() + 
  geom_point(data=dat2, aes(x=DO_mgL, y=Turb_NTU, fill=Chlor), pch=21, size=4) +
  scale_fill_viridis() + 
  facet_grid(Pond_Slough~.)

```

## Spatial Maps

```{r makeSpatial}

dat_sf <- st_as_sf(dat2,
                   coords = c("Long", "Lat"),
                   remove = F, 
                   crs = 4326) # add projection

```

```{r MapViewplot}

mapview(dat_sf)

```

```{r geom_sf}

ggplot() + 
  geom_sf(data=dat_sf, aes(fill=DO_sat), pch=21, size=4) +
  scale_fill_viridis() + theme_minimal()

```


